When, why, and how to avoid, prefer, or tolerate inbreeding

### 2018-03-04 presentation/inbreeding.Rmd
More slides up to model 2

<style type="text/css">
body p {
  color: #000000;
}
</style>

@@

**When should we predict selection for a alleles causing pre-copulatory inbreeding avoidance or inbreeding preference? [1]**

@@

- Selection against deleterious recessives
- Dominance coefficient of deleterious recessives
- Direct selection against inbreeding avoidance
- Offspring produced per female
- Maximum mates per male
- Immigrants per generation
- Female and male carrying capacity
- Allele mutation rate (forward & back)
- Effect strength of inbreeding avoidance allele

</div>
<span style="font-size: 18px;">
1. Duthie, AB & JM Reid (2016) *Am. Nat.* [**188**:651-667](https://bradduthie.github.io/Duthie_Reid2016_AppendixB.pdf). From [Appendix B](https://bradduthie.github.io/Duthie_Reid2016_AppendixB.pdf)
</span>

@@

$\delta \approx 0.22$ for full siblings

$\delta \approx 0.12$ for half siblings

$\delta \approx 0.06$ for first cousins

</div>
<span style="font-size: 18px;">
1. Duthie, AB & JM Reid (2016) *Am. Nat.* [**188**:651-667](https://bradduthie.github.io/Duthie_Reid2016_AppendixB.pdf). From [Appendix B](https://bradduthie.github.io/Duthie_Reid2016_AppendixB.pdf)
</span>

@@

</div>
<span style="font-size: 18px;">
1. Duthie, AB & JM Reid (2016) *Am. Nat.* [**188**:651-667](https://bradduthie.github.io/Duthie_Reid2016_AppendixB.pdf). From [Appendix B](https://bradduthie.github.io/Duthie_Reid2016_AppendixB.pdf)
</span>

@@

**Selection for inbreeding avoidance increases when the *effect* of the allele is strengthened (i.e., relatives are avoided at higher probability)**

</div>
<span style="font-size: 18px;">
1. Duthie, AB & JM Reid (2016) *Am. Nat.* [**188**:651-667](https://bradduthie.github.io/Duthie_Reid2016_AppendixB.pdf). From [Appendix B](https://bradduthie.github.io/Duthie_Reid2016_AppendixB.pdf)
</span>

@@

![](images/M_A.png)

@@

Selection for inbreeding avoidance increases when the *effect* of the allele is strengthened (i.e., relatives are avoided at higher probability)

**Selection for inbreeding avoidance comparable when males rather than females chose mates** (*Note:* Males were mate-limited)

</div>
<span style="font-size: 18px;">
1. Duthie, AB & JM Reid (2016) *Am. Nat.* [**188**:651-667](https://bradduthie.github.io/Duthie_Reid2016_AppendixB.pdf). From [Appendix B](https://bradduthie.github.io/Duthie_Reid2016_AppendixB.pdf)
</span>

## What to do

Add effective population size plot to females before showing males

After briefly showing males, then move on to showing inbreeding preference

Note that inbreeding preference is predicted for females, not males

Move on to different genetic architecture where inbreeding avoidance/preference in same sim

>>>>

<style type="text/css">
body p {
  color: #000000;
}
ul {
  color: black;
}
ol {
  color: black;
}
a {
    color: darkblue;
    text-decoration: none
}
</style>

@@

**When should we predict selection for alleles causing pre-copulatory inbreeding avoidance or inbreeding preference? [1]**

@@

1. Selection against deleterious recessives
2. Direct selection against inbreeding avoidance
3. Kin discrimination ability (all vs. only sibs)
4. Effect strength of inbreeding avoidance allele
5. Female and male carrying capacity
6. Dominance coefficient of deleterious recessives
7. Offspring produced per female
8. Maximum mates per male
9. Immigrants per generation
10. Allele mutation rate (forward & back)

</div>
<span style="font-size: 18px;">
Duthie, AB & JM Reid (2016) *Am. Nat.* [**188**:651-667](https://bradduthie.github.io/Duthie_Reid2016_AppendixB.pdf). From [Appendix B](https://bradduthie.github.io/Duthie_Reid2016_AppendixB.pdf)
</span>

1. [Selection against deleterious recessives](https://bradduthie.github.io/Duthie_Reid2016.pdf#page=7)
2. [Direct selection against inbreeding avoidance](https://bradduthie.github.io/Duthie_Reid2016.pdf#page=7)
3. [Kin discrimination ability (all vs. only sibs)](https://bradduthie.github.io/Duthie_Reid2016_SM.pdf#page=11)
4. [Effect strength of inbreeding avoidance allele](https://bradduthie.github.io/Duthie_Reid2016_SM.pdf#page=19)
5. [Female and male carrying capacity](https://bradduthie.github.io/Duthie_Reid2016_SM.pdf#page=17)
6. Dominance coefficient of deleterious recessives
7. Offspring produced per female
8. Maximum mates per male
9. Immigrants per generation
10. Allele mutation rate (forward & back)

</div>
<span style="font-size: 18px;">
Duthie, AB & JM Reid (2016) *Am. Nat.* [**188**:651-667](https://bradduthie.github.io/Duthie_Reid2016_AppendixB.pdf). From [Appendix B](https://bradduthie.github.io/Duthie_Reid2016_AppendixB.pdf)
</span>

![](images/neutrals.png)

1. [Selection against deleterious recessives](https://bradduthie.github.io/Duthie_Reid2016.pdf#page=7)
2. [Direct selection against inbreeding avoidance](https://bradduthie.github.io/Duthie_Reid2016.pdf#page=7)
3. [Kin discrimination ability (all vs. only sibs)](https://bradduthie.github.io/Duthie_Reid2016_SM.pdf#page=11)
4. [Effect strength of inbreeding avoidance allele](https://bradduthie.github.io/Duthie_Reid2016_SM.pdf#page=19)
5. [Female and male carrying capacity](https://bradduthie.github.io/Duthie_Reid2016_SM.pdf#page=17)
6. Dominance coefficient of deleterious recessives
7. Offspring produced per female
8. Maximum mates per male
9. Immigrants per generation
10. Allele mutation rate (forward & back)

</div>
<span style="font-size: 18px;">
Duthie, AB & JM Reid (2016) *Am. Nat.* [**188**:651-667](https://bradduthie.github.io/Duthie_Reid2016_AppendixA.pdf). From [Appendix A](https://bradduthie.github.io/Duthie_Reid2016_AppendixA.pdf)
</span>

@@

$\delta \approx 0.22$ for offspring of full siblings

$\delta \approx 0.12$ for offspring of half siblings

$\delta \approx 0.06$ for offspring of first cousins

</div>
<span style="font-size: 18px;">
Duthie, AB & JM Reid (2016) *Am. Nat.* [**188**:651-667](https://bradduthie.github.io/Duthie_Reid2016).
</span>

@@

</div>
<span style="font-size: 18px;">
Duthie, AB & JM Reid (2016) *Am. Nat.* [**188**:651-667](https://bradduthie.github.io/Duthie_Reid2016.pdf).
</span>

![](images/F_A_sib.png)

**Selection for inbreeding avoidance was limited when females could only identify siblings as kin.**

</div>
<span style="font-size: 18px;">
Duthie, AB & JM Reid (2016) *Am. Nat.* [**188**:651-667](https://bradduthie.github.io/Duthie_Reid2016.pdf). 
</span>

@@

Selection for inbreeding avoidance was limited when females could only identify siblings as kin.

**Selection for inbreeding avoidance increases when the *effect* of the allele is strengthened (i.e., relatives are avoided at higher probability).**

</div>
<span style="font-size: 18px;">
Duthie, AB & JM Reid (2016) *Am. Nat.* [**188**:651-667](https://bradduthie.github.io/Duthie_Reid2016.pdf). 
</span>

@@

![](images/F_A_Ne.png)

@@

Selection for inbreeding avoidance was limited when females could only identify siblings as kin.

Selection for inbreeding avoidance increases when the *effect* of the allele is strengthened (i.e., relatives are avoided at higher probability).

**Selection was stronger at *lower* N<sub>e</sub> due to *increased* inbreeding risk.**

</div>
<span style="font-size: 18px;">
Duthie, AB & JM Reid (2016) *Am. Nat.* [**188**:651-667](https://bradduthie.github.io/Duthie_Reid2016.pdf). 
</span>

## Mutant allele underlying inbreeding preference

![](images/F_P.png)

1000 loci with deleterious recessive alleles (homozygotes 0.4\% survival cost)

Inbreeding depression slope $\beta \approx 0.5$

$\delta \approx 0.12$ for offspring of full siblings

$\delta \approx 0.06$ for offspring of half siblings

$\delta \approx 0.03$ for offspring of first cousins

</div>
<span style="font-size: 18px;">
Duthie, AB & JM Reid (2016) *Am. Nat.* [**188**:651-667](https://bradduthie.github.io/Duthie_Reid2016).
</span>

## Population genetic model conclusions

**Overall, selection for a mutant allele underlying inbreeding avoidance:**

1. **Increased** given [stronger selection against deleterious recessives](https://bradduthie.github.io/Duthie_Reid2016.pdf#page=7)
2. **Decreased** given [direct selection against inbreeding avoidance](https://bradduthie.github.io/Duthie_Reid2016.pdf#page=7)
3. **Decreased** given restricted [kin discrimination ability (all vs. only sibs)](https://bradduthie.github.io/Duthie_Reid2016_SM.pdf#page=11)
4. **Increased** given an increasing [effect strength of inbreeding avoidance allele](https://bradduthie.github.io/Duthie_Reid2016_SM.pdf#page=19)
5. **Decreased** given increased [female and male carrying capacity](https://bradduthie.github.io/Duthie_Reid2016_SM.pdf#page=17) (and consequent increase in effective population size)

</div>
<span style="font-size: 18px;">
Duthie, AB & JM Reid (2016) *Am. Nat.* [**188**:651-667](https://bradduthie.github.io/Duthie_Reid2016).
</span>

## Population genetic model conclusions

**Additional conclusions:**

- [Comparable selection](https://bradduthie.github.io/Duthie_Reid2016_SM.pdf\page=11) for inbreeding avoidance alleles was observed when males chose (*note*, males were mate limited when choosing).
- Weak inbreeding depression caused selection for inbreeding preference when *females* chose, but not when *males* chose mates.
- [No evidence of purging](https://bradduthie.github.io/Duthie_Reid2016_SM.pdf#page=5) in simulations with mutant alleles underlying inbreeding avoidance or preference.

<br><br><br><br><br><br>

</div>
<span style="font-size: 18px;">
Duthie, AB & JM Reid (2016) *Am. Nat.* [**188**:651-667](https://bradduthie.github.io/Duthie_Reid2016).
</span>

## Multi-locus model with infinite alleles

![](images/beta_slopes.png)

Multi-locus genetic architecture:

- Fixed inbreeding depression for individual *f*
- Allele values at 10 loci increase (positive) or decrease (negative) inbreeding
- Allele values sum additively to determine overall inbreeding behaviour phenotype

<span style="font-size: 18px;">
Duthie, AB et al. (2016) *Evolution* [**70**:1927-1943](https://bradduthie.github.io/DuthieEtAl2016b.pdf).
</span>

## Multi-locus model with infinite alleles

![](images/IA_quant.png)

Multi-locus genetic architecture:

- Fixed inbreeding depression for individual *f*
- Allele values at 10 loci increase (positive) or decrease (negative) inbreeding
- Allele values sum additively to determine overall inbreeding behaviour phenotype

<span style="font-size: 18px;">
Duthie, AB et al. (2016) *Evolution* [**70**:1927-1943](https://bradduthie.github.io/DuthieEtAl2016b.pdf).
</span>

## Inbreeding adjustment through multiple mating {.smaller}
